Multi-fluids for Representing Sub-\change[HW]{filter}{grid}-scale Convection

### 2020-07-17 submit3/manuscript.tex
Included Will's comments

\begin{abstract}
\add[HW]{Traditional parameterizations of convection are a large source of error in weather and climate prediction models and the assumptions behind them are worse as resolution increases.}
Multi-fluid modeling is a promising new method of representing sub- and near-grid-scale convection allowing for \remove[HW]{consistent and accurate
treatment of} net mass transport by convection and \remove[HW]{of} non-equilibrium dynamics. The air is partitioned into two or more
fluids which may represent, for example, updrafts and the \change[HW]{stable}{non-updraft} environment. Each fluid has its own velocity, temperature and constituents with separate equations of motion.

@@

Next the conditional averaging of terms in the momentum (\ref{eq:singleMom}) and buoyancy (\ref{eq:singleb}) equations is considered. \add[HW]{The filter scale is assumed to be equal to the grid scale but the derivation does not rely on this.} Conditionally averaging the advection terms leads to sub-filter-scale fluxes, $F_{SF}$, since the velocity and buoyancy are not uniform within each fluid at the filter scale:
\begin{eqnarray}
\overline{I_i \nabla \cdot (\mathbf{u} \mathbf{u})} &=& \nabla \cdot (\sigma_i \mathbf{u}_i  \mathbf{u}_i) + F_\text{SF}^{\mathbf{u}_i} \\
\overline{I_i \nabla (\mathbf{u} b)} &=&  \nabla \cdot (\sigma_i \mathbf{u}_i b_i) + F_\text{SF}^{b_i}
\label{eq:filterAdvection}
\end{eqnarray}
The sub-filter-scale fluxes are represented by constant viscosity diffusion terms for each fluid. The diffusion terms are written so that, in the absence of transfers between fluids and if both fluids are initialized to be identical, the fluids remain identical and $\sigma_i$ acts as a passive tracer:
\begin{eqnarray}
\overline{I_i \nu \nabla^2 \mathbf{u}} &\approx& \sigma_i \nu \nabla^2 \mathbf{u}_i \\
\overline{I_i \alpha \nabla^2 b} &\approx& \sigma_i \alpha \nabla^2b_i .
\end{eqnarray} 
%
The \add[HW]{Boussinesq} pressure \add[HW]{potetials} within each fluid are defined as:
\begin{equation}
\sigma_i P_i = \overline{I_i P}.
\end{equation}
The pressure gradient can then be decomposed as:
\begin{equation}
\overline{I_i \nabla P} = \sigma_i \nabla P_i + 
    \left[ P_i\nabla\sigma_i - \overline{P\nabla I_i} \right]
\end{equation}
The terms in square brackets are the difference between the pressure force pushing outward from the fluid calculated using resolved variables and the full resolved version of the same force, both of which are related to drag. They will be represented by a drag, $\mathbf{D}_{ij}$. However this drag is likely to be smaller than the drag that is used, for example, in multi-phase modeling \cite<e.g.>{GBB+07} as it is representing the difference between two representations of drag, rather than the whole drag. Sensitivity to $\mathbf{D}_{ij}$ will be investigated in section \ref{sec:results}.

@@

The instabilities reported by \citeA{WM19} showed growing divergence
in a fluid with vanishing volume fraction. For single-fluid equations, divergence would lead to pressure anomalies which would act as a negative feedback on the divergence. However for multi-fluids with just one pressure, if one of the volume fractions is too small then it cannot force pressure anomalies big enough to stabilize the divergence. Use of a separate pressure for each fluid is therefore proposed to ensure stability.

@@

In order to derive a suitable form of \change[HW]{parameterization}{model} for $p_i$ we will first consider the \citeA{BN86} model of multi-phase flow.

@@

\citeA{TEB19} stabilized the multi-fluid equations by adding a diffusion
term $\nu\partial w^{2}/\partial z^{2}$ to the RHS of the $w$ equation.
If a fluid perturbation pressure is set to $p_i=-\nu\partial w/\partial z$
then this is equivalent to the diffusion term in one dimension. However adding an artificial diffusion term will harm three dimensional simulations of
resolved convection in the free troposphere whereas adding a fluid
perturbation pressure will not. Inspired by both \citeA{BN86} and
\citeA{TEB19}
and obeying the constraints given above, the pressure in each fluid is chosen to be:
\begin{equation}
p_{i}=
-\underbrace{\gamma\nabla\cdot\mathbf{u}_{i}}_{\text{stabilization}}
+
\underbrace{\gamma\sum_j \sigma_j\nabla\cdot\mathbf{u}_j}_{\text{correction}}
\label{eq:Pi_div}
\end{equation}
where $\gamma$ is a compressibility with units $\text{m}^2 \text{s}^{-1}$. The ``stabilization'' term reduces to that of \protect{\citeA{TEB19}} in one dimension and the ``correction'' term ensures that $\sum_i \sigma_i p_i =0$. It should be noted that this parameterization destroys \add[HW]{exact} energy conservation. \add[HW]{In future work solving the compressible equations we plan to re-insert the energy lost due to this stabilization as turbulent kinetic energy and heat.}

@@

Using entrainment and detrainment set by divergence ($S_{ij}=-\nabla\cdot\mathbf{u}_{i}$)
and pressure difference between the fluids controlled by divergence
($P_{i}=P+\gamma\sum_j \sigma_j \nabla\cdot\mathbf{u}_j-\gamma\nabla\cdot\mathbf{u}_{i}$), the simulations are more realistic (bottom three rows of figure \ref{fig:multiFluidBubble_500_1}).
The first value, $\gamma=2.3\times10^{4}\text{m}^{2}\text{s}^{-1}$, is set from equation
(\ref{eq:gammaDimAnal}) using length scale equal to the initial bubble
radius and a buoyancy scale equal to the initial maximum buoyancy. This makes
the velocity profile too smooth and the updraft too weak so the
bubble does not rise enough. It is reasonable to consider smaller values of $\gamma$ because the buoyancy does not retain its initial maximum for very long. Using $\gamma=10^{4}\text{m}^{2}\text{s}^{-1}$ gives more accurate solutions whereas using $\gamma=4\times10^{3}\text{m}^{2}\text{s}^{-1}$ leads to the leading edge of the bubble rising too quickly and does not fully remove the discontinuity in the updraft velocity at the leading edge of the bubble. The pressure differences between the fluids have some similarity with the resolved simulation but none of the values of $\gamma$ reproduce the pressure dip at the trailing edge of the bubble. $\gamma=10^{4}\text{m}^{2}\text{s}^{-1}$ gives the closest approximation to the resolved simulation.

@@

The properties of the fluid that are transferred from one fluid to another are unknowns in the multi-fluid equations and these require closure. An obvious
first guess is that fluid takes its mean buoyancy and velocity
with it when it is transferred.
Improvements on this assumption will depend on knowledge of sub-filter-scale variability. However since conditional averaging is based on the sign of $w$, it is consistent to assume that the fluid transferred has $w=0$ ($w_{ij}^T=0$).
Three models for the buoyancy of the fluid transferred 
are tested. The most successful
is to assume that the fluid takes its mean buoyancy with it
($b_{ij}^T=b_i$).
Assuming that the buoyancy of the transferred fluid is an average of the buoyancy of both fluids ($b_{ij}^T=\frac{1}{2}(b_i+b_j)$) can lead to unrealistically large or small, unbounded buoyancies. 
Assuming that one fluid has negative buoyancy and the other positive buoyancy and fluid of zero buoyancy is transferred ($b_{ij}^T=0$) also leads to unbounded buoyancy.

>>>>

\begin{abstract}
\add[HW]{Traditional parameterizations of convection are a large source of error in weather and climate prediction models and the assumptions behind them become worse as resolution increases.}
Multi-fluid modeling is a promising new method of representing sub- and near-grid-scale convection allowing for \remove[HW]{consistent and accurate
treatment of} net mass transport by convection and \remove[HW]{of} non-equilibrium dynamics. The air is partitioned into two or more
fluids which may represent, for example, updrafts and the \change[HW]{stable}{non-updraft} environment. Each fluid has its own velocity, temperature and constituents with separate equations of motion.

@@

Next the conditional averaging of terms in the momentum (\ref{eq:singleMom}) and buoyancy (\ref{eq:singleb}) equations is considered. \add[HW]{The filter scale is assumed to be equal to the grid scale but the derivation does not rely on this.} Conditionally averaging the advection terms leads to sub-filter-scale fluxes, $F_{SF}$, since the velocity and buoyancy are not uniform within each fluid at the filter scale:
\begin{eqnarray}
\nabla \cdot \overline{I_i \mathbf{u} \mathbf{u}} &=& \nabla \cdot (\sigma_i \mathbf{u}_i  \mathbf{u}_i) + F_\text{SF}^{\mathbf{u}_i} \\
\nabla \cdot \overline{I_i \mathbf{u} b} &=&  \nabla \cdot (\sigma_i \mathbf{u}_i b_i) + F_\text{SF}^{b_i}
\label{eq:filterAdvection}
\end{eqnarray}
The sub-filter-scale fluxes are represented by constant viscosity diffusion terms for each fluid. The diffusion terms are written so that, in the absence of transfers between fluids and if both fluids are initialized to be identical, the fluids remain identical and $\sigma_i$ acts as a passive tracer:
\begin{eqnarray}
\nu \nabla^2 \overline{I_i \mathbf{u}} &\approx& \sigma_i \nu \nabla^2 \mathbf{u}_i \\
\alpha \nabla^2 \overline{I_i b} &\approx& \sigma_i \alpha \nabla^2b_i .
\end{eqnarray} 
%
The \add[HW]{Boussinesq} pressure \add[HW]{potentials} within each fluid are defined as:
\begin{equation}
\sigma_i P_i = \overline{I_i P}.
\end{equation}
The pressure gradient can then be decomposed as:
\begin{equation}
\overline{I_i \nabla P} = \sigma_i \nabla P_i + 
    \left[ P_i\nabla\sigma_i - \overline{P\nabla I_i} \right]
\end{equation}
The terms in square brackets are the difference between the pressure force pushing outward from the fluid calculated using \change[HW]{resolved variables and the full resolved version}{the filtered variables and the fully resolved version} of the same force, both of which are related to drag. They will be represented by a drag, $\mathbf{D}_{ij}$. However this drag is likely to be smaller than the drag that is used, for example, in multi-phase modeling \cite<e.g.>{GBB+07} as it is representing the difference between two representations of drag, rather than the whole drag. Sensitivity to $\mathbf{D}_{ij}$ will be investigated in section \ref{sec:results}.

@@

The instabilities reported by \citeA{WM19} showed growing divergence
in a fluid with vanishing volume fraction. For single-fluid equations, divergence would lead to pressure anomalies which would act as a negative feedback on the divergence. \change[HW]{However for multi-fluids with just one pressure, if one of the volume fractions is too small then it} {In the multi-fluid case with just one pressure, fluids with a sufficiently small volume fraction} cannot force pressure anomalies big enough to stabilize the divergence. Use of a separate pressure for each fluid is therefore proposed to ensure stability.

@@

In order to derive a suitable \change[HW]{form of parameterization}{model} for $p_i$ we will first consider the \citeA{BN86} model of multi-phase flow.

@@

\citeA{TEB19} stabilized the multi-fluid equations by adding a diffusion
term $\nu\partial w^{2}/\partial z^{2}$ to the RHS of the $w$ equation.
If a fluid perturbation pressure is set to $p_i=-\nu\partial w/\partial z$
then this is equivalent to the diffusion term in one dimension. However adding an artificial diffusion term will harm three dimensional simulations of
resolved convection in the free troposphere whereas adding a fluid
perturbation pressure will not. Inspired by both \citeA{BN86} and
\citeA{TEB19}
and obeying the constraints given above, the pressure in each fluid is chosen to be:
\begin{equation}
p_{i}=
-\underbrace{\gamma\nabla\cdot\mathbf{u}_{i}}_{\text{stabilization}}
+
\underbrace{\gamma\sum_j \sigma_j\nabla\cdot\mathbf{u}_j}_{\text{correction}}
\label{eq:Pi_div}
\end{equation}
where $\gamma$ is a compressibility with units $\text{m}^2 \text{s}^{-1}$. The ``stabilization'' term reduces to that of \protect{\citeA{TEB19}} in one dimension and the ``correction'' term ensures that $\sum_i \sigma_i p_i =0$. It should be noted that this parameterization destroys \add[HW]{exact} energy conservation. \add[HW]{In future work solving the compressible equations, the energy lost due to this stabilization will be re-inserted as turbulent kinetic energy or heat.}

@@

Using entrainment and detrainment set by divergence ($S_{ij}=-\nabla\cdot\mathbf{u}_{i}$)
and pressure difference between the fluids controlled by divergence
($P_{i}=P-\gamma\nabla\cdot\mathbf{u}_{i}+\gamma\sum_j \sigma_j \nabla\cdot\mathbf{u}_j$), the simulations are more realistic (bottom three rows of figure \ref{fig:multiFluidBubble_500_1}).
The first value, $\gamma=2.3\times10^{4}\text{m}^{2}\text{s}^{-1}$, is set from equation
(\ref{eq:gammaDimAnal}) using length scale equal to the initial bubble
radius and a buoyancy scale equal to the initial maximum buoyancy. This makes
the velocity profile too smooth and the updraft too weak so the
bubble does not rise enough. It is reasonable to consider smaller values of $\gamma$ because the buoyancy does not retain its initial maximum for very long. Using $\gamma=10^{4}\text{m}^{2}\text{s}^{-1}$ gives more accurate solutions whereas using $\gamma=4\times10^{3}\text{m}^{2}\text{s}^{-1}$ leads to the leading edge of the bubble rising too quickly and does not fully remove the discontinuity in the updraft velocity at the leading edge of the bubble. The pressure differences between the fluids have some similarity with the resolved simulation but none of the values of $\gamma$ reproduce the pressure dip at the trailing edge of the bubble. $\gamma=10^{4}\text{m}^{2}\text{s}^{-1}$ gives the closest approximation to the resolved simulation.

@@

The properties of the fluid that are transferred from one fluid to another are unknowns in the multi-fluid equations and these require closure. An obvious
first guess is that fluid takes its mean buoyancy and velocity
with it when it is transferred.
Improvements on this assumption will depend on knowledge of sub-filter-scale variability. However since conditional averaging is based on the sign of $w$, it is consistent to assume that the fluid transferred has zero vertical velocity ($w_{ij}^T=0$).
Three models for the buoyancy of the fluid transferred 
are tested. The most successful
is to assume that the fluid takes its mean buoyancy with it
($b_{ij}^T=b_i$).
Assuming that the buoyancy of the transferred fluid is an average of the buoyancy of both fluids $\bigl(b_{ij}^T=\frac{1}{2}(b_i+b_j)\bigr)$ can lead to unrealistically large or small, unbounded buoyancies. 
Assuming that one fluid has negative buoyancy and the other positive buoyancy and fluid of zero buoyancy is transferred ($b_{ij}^T=0$) also leads to unbounded buoyancy.
